---
title: "Yara Colombia land use model (Draft)"
author: "Ben Schiek"
date: "March 7, 2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Yield $y$ is modeled as a Cobb-Douglas function of agricultural inputs $\mathbf{x}$, technology (input elasticities of yield) $\boldsymbol{\alpha}$, an exogenous scale factor $y_0$, and exogenous factors $\mathbf{k}$ with their corresponding elasticities $\boldsymbol{\beta}$.

\begin{equation}
y(\mathbf{x}; \mathbf{k}, \boldsymbol{\alpha}, \boldsymbol{\beta}) = y_0 \Pi_i x_i^{\alpha_i} \Pi_i k_i^{\beta_i}
\end{equation}

The Cobb-Douglas function is famous for its good empirical fit across a wide range of contexts and scales. In Appendix \ref{sec:} I show how the Cobb-Douglas function can be derived from the law of diminishing marginal returns.

In farm yield studies, the exogenous factors $\mathbf{k}$ typically contain demographic variables such as age and education, as well as agroclimatic variables such as soil quality, temperature, and precipitation. These variables can vary considerably across a farm population. Analysis is said to be "ceteris paribus" when these variables are held constant at particular values.

The scale factor $y_0$ is often interpreted as "total factor productivity", that is, overall productivity with respect to all variables []. This implies awkward units of, for example, output per unit volume man-hour kilogram, in the case where input variables are pesticide, labor, and fertilizer, respectively. A more straightforward interpretation suggests itself considering that ...

ceteris paribus unit input yield...basal yield as a shorthand.

Optimal (or "economic") yield $y^*$ is defined as yield evaluated at the optimal input levels $\mathbf{x}^*$. The optimal input levels are those that maximize net benefit subject to a budget constraint. Given the Cobb-Douglas functional form, optimal yield works out to

\begin{equation}
y^*(y_0, \chi, \kappa, \gamma) = y(\mathbf{x}^*) = (y_0 \chi \gamma \lambda^{-h})^(\frac{1}{1 - h})
\end{equation}

Where

\begin{equation}
h = \Sum_i \alpha_i
\end{equation}

is the yield degree of homogeneity, and

\begin{equation}
\gamma = \frac{P^h}{\Pi_i w_i^{\alpha_i}}
\end{equation}

\begin{equation}
\chi = \Pi_i \alpha_i^{\alpha_i}
\end{equation}

\begin{equation}
\kappa = \Pi_i k_i^{\beta_i}
\end{equation}

Input and output prices vary by department in Colombia. This sub-national variation can be captured in our model using the price data available from DANE and/or the Colombian Ministry of Agriculture.

The following table summarizes the input variables required by the yield model. The basal yield $y_0$, input elasticities $\alpha_i$, and exogenous control elasticities of yield $\beta_i$ are ideally fit to data, but can also be deduced from expert consultation and/or literature if need be.

The distance to market data can be included among the $k_i$.

```{r}


var_vec <- c("P", "w_i", "k_i", "y_0", "U+03B1", "\beta_i")

varExpl_vec <- c("output price",
                 "input price",
                 "demographic and/or agroclimatic control variables",
                 "basal yield",
                 "input elasticities of yield",
                 "control variable elasticities of yield")

avail_from <- c("https://www.agronet.gov.co/estadistica/paginas/home.aspx",
"https://www.dane.gov.co/index.php/en/statistics-by-topic/agriculture")

#"https://www.dane.gov.co/index.php/en/statistics-by-topic-1/agricultural-sector/national-agricultural-survey-ena"


df_table <- data.frame(Variable = var_vec,
                       avail_from)


```






# The land use model

A model of how much area is allocated to each agricultural commodity can be built from the log-linear empirical relation between the mean value of agricultural production (VAP) and the standard deviation of VAP. The mean and standard deviation are taken over the years 1961-2018.

```{r, fig.show = "hold", fig.width = 5, fig.height=4, fig.align="left", fig.cap="\\label{fig:foodSec}A log-linear empirical relation exists between mean VAP and standard deviation of VAP.", echo = FALSE}

this_folder <- "C:/Users/bensc/OneDrive/Documents/Data/FAO Data/"
this_file <- "Value_of_Production_E_All_Data.csv"
this_filepath <- paste0(this_folder, this_file)
df_raw <- read.csv(this_filepath, stringsAsFactors = F)
df_raw$Area.Code <- NULL
df_raw$Item.Code <- NULL
df_raw$Element.Code <- NULL
colnames(df_raw) <- gsub("Y", "", colnames(df_raw))
ind_rm <- grep("F", colnames(df_raw))
df_raw <- df_raw[, -ind_rm]
df_raw <- df_raw %>% gather(Year, Value, `1961`:`2018`)
these_areas <- c("World", "Colombia", "South America", "Central America")
df_vap <- subset(df_raw, Area %in% these_areas)
df_vap$Item[grep("Vegetables and Fruit Primary", df_vap$Item)] <- "Fruit & Veg."
df_vap$Item[grep("Oilcrops", df_vap$Item)] <- "Oilcrops"

pulses_vec <- c("Beans, dry", "Bambara beans", "Lentils", "Chick peas", "Peas, dry",
                "Cow peas, dry", "Pigeon peas", "Pulses nes")

df_pulses <- subset(df_vap, Item %in% pulses_vec)
df_pulses <- df_pulses %>% group_by(Area, Element, Unit, Year) %>%
  summarise(Value = sum(Value, na.rm = T)) %>% as.data.frame()
df_pulses$Item <- "Pulses, Total"
df_pulses <- df_pulses[, colnames(df_vap)]
df_vap <- as.data.frame(rbind(df_vap, df_pulses))
colnames(df_vap)[ncol(df_vap)] <- "VAP"
#----------------------------------------------------------------------------
these_areas <- c("World", "Colombia", "South America", "Central America")
df_vap <- subset(df_vap, Area %in% these_areas &
                   Element == "Gross Production Value (constant 2014-2016 1000 I$)")
# "Gross Production Value (constant 2014-2016 million US$)"
#unique(df_vap$Element)
df_mu <- df_vap %>% group_by(Area, Item) %>%
  summarise(mu = mean(VAP, na.rm = T)) %>% as.data.frame()
df_sd <- df_vap %>% group_by(Area, Item) %>%
  summarise(sd = sd(VAP, na.rm = T)) %>% as.data.frame()
df_plot <- merge(df_mu, df_sd, by = c("Area", "Item"))
df_plot <- subset(df_plot, Area %in% these_areas)
df_plot$`logged mean VAP` <- log(df_plot$mu)
df_plot$`logged stand. dev. VAP` <- log(df_plot$sd)

gg <- ggplot(df_plot, aes(x = `logged stand. dev. VAP`, y = `logged mean VAP`))
gg <- gg + geom_point()
#gg <- gg + scale_x_log10() + scale_y_log10()
gg <- gg + facet_wrap(~Area, scales = "free")
gg

```

The regression for Colombia is:

```{r}

df_mod <- subset(df_plot, Area == "Colombia")
mod <- lm(lmu ~ lsd, df_mod)
summary(mod)
yint <- coefficients(mod)[1]
slope <- coefficients(mod)[2]

```

The regression equation is:

\begin{equation}
\ln(\mu_i) \approx `r slope` \ln(\sigma_i) + `r yint`
\end{equation}

Where $\mu_i$ and $\sigma_i$ are the mean VAP and standard deviation of VAP, respectively, defined as follows.

\begin{equation}
\mu_i = a_i E[P_i y_i^*]
\sigma = a_i \sqrt{Var[P_i y_i^*]}
\end{equation}

Where $a_i$ is the harvested area of commodity $i$.

The fitted equation can then be solved for harvested area $a_i$.

\begin{equation}
a_i \approx \left( e^{`r yint`} \frac{\sigma^{`r slope`}}{P_i y_i^*} \right)^{\frac{1}{1 - `r slope`}}
\end{equation}

This gives us a model of area as a function of yield and price risk $\sigma_i$, price $P_i$, yield $y^*$ (which is a function of many other things, as derived above), and the fitted parameters `r slope` and `r yint`.

$\rightarrow$ An increase in the expected revenue from commodity $i$, whether through an increase in output price or increase in yield (from technology adoption, for example), results in lower area allocated to that commodity.

$\rightarrow$ A decrease in the price and yield risk $\sigma_i$ of commodity $i$ also reduces area allocated to the commodity.

$\Rightarrow$ New technology that reduces yield risk and/or increases yield reduces area (for given prices). Uptake of such technology may thus be considered sustainable intestification.

# Modeling new technology release and uptake

New technology can be modeled either as a change in one of the $\alpha_i$ or as a change in the basal yield (a.k.a. "total factor productivity") $y_0$.

If the change in technology comes in the form of a more efficient input which leaves the other input technology unaltered, then this can be modeled as an increase in the corresponding yield response $\alpha_i$. Assuming that only those who stand to increase their net revenue will adopt, the following adoption condition can be derived (details in Appendix).

\begin{equation}
y^* > y_0 (w_i / \alpha_i e^{\eta(\mathbf{w}, \alpha_i) - \eta(P_i, \alpha_i)})^h \chi \kappa \gamma
\end{equation}

That is, only farms (or pixels) with a baseline yield higher than this will find it profitable to adopt. Farms with a baseline yield below this would incur a loss in net benefit if they were to adopt.

If the change in technology comes in the form of improved genetic material that effectively improves yield response to several inputs at once, it is more appropriate to model this as a change in the basal yield $y_0$. In this case, the adoption condition works out to

\begin{equation}
\frac{\partial \ln (\gamma)}{\partial \ln (y_0)} > -1
\end{equation}

For this study, it seems to me that the first technology model applies. I.e. we model Yara fertilizer adoption as an increase in the $\alpha_i$ corresponding to fertilizer. In other words, we increase the fertilizer elasticity of yield.

# Commodity focus

The following commodities account for ~90% of total cropland in Colombia as of 2017. "Total cropland" comes from the CCI LC remote sensing data (herbaceous crops), downloaded from FAOSTAT.

```{r}

this_file <- "Environment_LandCover_E_All_Data.csv"
this_filepath <- paste0(this_folder, this_file)
df_raw <- read.csv(this_filepath, stringsAsFactors = F)
rm_cols <- grep("Code|F", colnames(df_raw))
df_raw <- df_raw[, -rm_cols]
colnames(df_raw) <- gsub("Y", "", colnames(df_raw))
df_raw <- df_raw %>% gather(Year, Value, `1992`:`2018`)
df_cover <- subset(df_raw, Element == "Area from CCI_LC")
df_cover$Element <- NULL 
df_cover$Value <- 1000 * df_cover$Value
df_cover$Unit <- NULL
df_cropCover <- subset(df_cover, Item == "Herbaceous crops")
df_grassCover <- subset(df_cover, Item == "Grassland")
df_cropCover$Item <- NULL
df_grassCover$Item <- NULL
colnames(df_cropCover)[ncol(df_cropCover)] <- "Total cropland"
colnames(df_grassCover)[ncol(df_grassCover)] <- "Total grassland"
#----------------------------------------------------------------------------
df_area <- subset(df_crop, Element == "Area harvested")
df_area$Element <- NULL
colnames(df_area)[ncol(df_area)] <- "Area harv"
df_area <- merge(df_area, df_cropCover, by = c("Area", "Year"))
df_area$`Area share` <- 100 * df_area$`Area harv` / df_area$`Total cropland`

df_q <- subset(df_area, Area == "Colombia" & Year == 2017)
df_q <- df_q[-grep(", Total|Rice Milled Eqv", df_q$Item), ]

ind_top <- which(df_q$`Area share` > 8)
df_q[ind_top, ]
sum(df_q$`Area share`[ind_top])

# q <- quantile(df_q$`Area share`, probs = 0.9, na.rm = T)
# ind_top <- which(df_q$`Area share` > q)
# df_q[ind_top, ]
# sum(df_q$`Area share`[ind_top])

df_table <- df_q


```

# Modeling GHG impacts

Can build off GHG conversion factors on page 12 of Bangladesh study.